---
title: "220307_fitness_data_analysis"
author: "Kyle Padoock"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/")
source('210304_microbe_functions.R')
```

##Summary
Soil microbiomes can provide plants with increased pest defenses. They also may alter fitness directly of the insects living in them. Here we have collected fitness data from WCR larvae and corn plants grown in the presence of cover-crop managed soils and traditionally managed soils. First let's look at the data.


```{r}
setwd("..")
fit.dat <- read_csv("Data/Fitness/210602_fitness_data.csv")
view(fit.dat)
```

To start, let's look at dry weight which likely would be most effected by microbial disruption and plant defenses in such a short bioassay.

In a previous analysis, we found that the dry weight residuals do not satisfy the assumptions of normality and were transformed. We square root transformed all values.


```{r}
#Remove uninfested controls
fit.dat.wcr <- fit.dat %>% dplyr::filter(colony != "CTRL")
fit.dat.wcr$block <- as.factor(fit.dat.wcr$block)

#Transform dry weight values
weight.tran <- fit.dat.wcr %>% 
  dplyr::select(treatment, corn_trait, colony, rep, block, weight_per_larvae) %>% 
  mutate(sqrt_wt = weight_per_larvae^(1/2)) 

table(weight.tran$treatment, weight.tran$corn_trait, weight.tran$colony)
# anova model
aov.fit.mod.sqrt <- aov(sqrt_wt ~ treatment * corn_trait * colony + block + block*corn_trait + block*colony, data = weight.tran)
shapiro.test(residuals(aov.fit.mod.sqrt))
summary(aov.fit.mod.sqrt)
fit.emm.treat <- emmeans(aov.fit.mod.sqrt, pairwise ~ treatment|colony)
plot(fit.emm.treat)
fit.emm.trait <- emmeans(aov.fit.mod.sqrt, pairwise ~ corn_trait|colony)
plot(fit.emm.trait)

TukeyHSD(aov.fit.mod.sqrt, which = "treatment:colony")
TukeyHSD(aov.fit.mod.sqrt, which = "treatment:corn_trait:colony")

fit.dat.wcr2 <- fit.dat.wcr %>% group_by(treatment, corn_trait, colony, block) %>% # calculate median values
  mutate(dw.med = ifelse(weight_per_larvae,median(weight_per_larvae, na.rm=TRUE), ifelse(weight_per_larvae == NA, NA))) 
fit.dat.wcr2$dat.dw.res<-abs(fit.dat.wcr2$weight_per_larvae-fit.dat.wcr2$dw.med) #calculate residual variance
levene.dat.aov.dw<-aov(dat.dw.res~treatment*corn_trait*colony*block,fit.dat.wcr2)
summary(levene.dat.aov.dw)
TukeyHSD(levene.dat.aov.dw)

fit.dat.wcr3 <- fit.dat.wcr %>% 
  dplyr::mutate(treatment2 = treatment) %>% 
  dplyr::mutate(corn2 = corn_trait) %>% 
  unite(treat_corn, treatment2, corn2, sep = '_')

fit.dat.wcr3 <- fit.dat.wcr3 %>% 
  dplyr::mutate(corn3 = corn_trait) %>% 
  dplyr::mutate(colony2 = colony) %>% 
  unite(corn_colony, corn3, colony2, sep = '_')

weight.tran2 <- fit.dat.wcr3 %>% 
  dplyr::select(treatment, treat_corn, corn_colony, colony, rep, block, weight_per_larvae) %>% 
  mutate(sqrt_wt = weight_per_larvae^(1/2)) 

aov.fit.mod.sqrt2 <- aov(sqrt_wt ~ treat_corn * colony + block + block*colony, data = weight.tran2)
shapiro.test(residuals(aov.fit.mod.sqrt2))
summary(aov.fit.mod.sqrt2)
fit.emm.treat_corn <- emmeans(aov.fit.mod.sqrt2, pairwise ~ treat_corn|colony)
plot(fit.emm.treat_corn)

aov.fit.mod.sqrt3 <- aov(sqrt_wt ~ treatment * corn_colony + block, data = weight.tran2)
shapiro.test(residuals(aov.fit.mod.sqrt3))
summary(aov.fit.mod.sqrt3)
fit.emm.corn_colony <- emmeans(aov.fit.mod.sqrt3, pairwise ~ corn_colony|treatment)
plot(fit.emm.corn_colony)

```

We see here there is a significant corn type x colony interaction. This is capturing the difference between res and susceptible on Bt. What's interesting is we also see a treatment x colony interaction. To further investigate this, we can separate the colonies and analyze them individually.

```{r}
fit.dat.wcr.res <- weight.tran %>% 
  filter(colony == "RES")

aov.fit.mod.res <- aov(sqrt_wt ~ treatment * corn_trait + block + block*corn_trait, data = fit.dat.wcr.res)
shapiro.test(residuals(aov.fit.mod.res))
summary(aov.fit.mod.res)
TukeyHSD(aov.fit.mod.res, which = "treatment")
TukeyHSD(aov.fit.mod.res, which = "treatment:corn_trait")

fit.dat.wcr.sus <- weight.tran %>% 
  filter(colony == "SUS")

aov.fit.mod.sus <- aov(sqrt_wt ~ treatment * corn_trait + block + block*corn_trait, data = fit.dat.wcr.sus)
shapiro.test(residuals(aov.fit.mod.sus))
summary(aov.fit.mod.sus)
TukeyHSD(aov.fit.mod.sus, which = "treatment")

```
Interesting, we find that resistant insects are slightly lighter when grown in traditionally managed soils while susceptible insects do slightly better in cover crop managed. We don't see this in the emmeans test, which is probably the most accurate way to analyze these data. The interaction is due to the resistant insects being effected by the soil treatment but not the susceptible. However, this is probably due to variation in the dry weight from averaging across Bt plants and non-bt plants. Maybe we should treat each combination as a unique group to see how it changes.

```{r}
# summary of average weight and standard error
sum.weight <- fit.dat.wcr %>% 
  dplyr::group_by(colony, corn_trait, treatment) %>% 
  dplyr::summarise(mean_weight = mean(weight_per_larvae, na.rm = TRUE),
                   sd = std.error(weight_per_larvae, na.rm = TRUE))

# generate figure with means, std. error and raw data
supp.labs <- c("Resistant", "Susceptible")
names(supp.labs) <- c("RES", "SUS")

weight.plot <- sum.weight %>% 
  ggplot(aes(x = corn_trait, y = mean_weight, colour= treatment)) +
  geom_point(position = position_dodge(0.5)) + 
  scale_color_manual(values = c("#C3BD10", "#137098"), labels = c("Conservation", "Traditional")) +
  geom_errorbar(aes(ymin=mean_weight-sd, ymax=mean_weight+sd), width=0.07, position = position_dodge(0.5)) +
  facet_wrap(~colony, scales = "free_y", labeller = labeller(colony = supp.labs))


weight.plot + geom_violin(aes(x = corn_trait, y = weight_per_larvae), data = fit.dat.wcr, position = position_dodge(0.5), alpha=0.5) + theme_classic() + xlab("Corn type") +ylab("Larval weight (g)") + labs(color = "Treatment") + 
  theme(strip.text.x = element_text(size = 20),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) 


ggsave("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/Data/Fitness/220502_dry_weight.pdf", dpi = 300, width = 9, height = 7)
ggsave("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/Data/Fitness/220427_dry_weight.tiff", dpi = 300, width = 9, height = 7)

fit.dat.wcr2 <- fit.dat.wcr%>%
   mutate(across(colony, factor, levels=c("SUS","RES")))

weight.plot2 <- 
  ggplot(aes(x = corn_trait, y = weight_per_larvae, fill= treatment), data = fit.dat.wcr2) +
  geom_violin() + scale_fill_manual(values = c("#C3BD10", "#137098"), labels = c("Cover crop no-till", "Traditional")) +
  geom_point(position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.2), alpha = .5) +
  geom_boxplot(width=0.1, position = position_dodge(width = .9)) + scale_x_discrete(labels = c("Bt corn", "non-Bt corn")) +
  facet_wrap(~colony, scales = "fixed",labeller = labeller(colony = supp.labs)) +
  theme_classic() + xlab("Corn type") +ylab("Average larval weight (g)") + labs(fill = "Treatment") +
  theme(
    strip.text.x = element_text(size = 26),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 20),
    axis.title.x = element_text(size = 24, margin = margin(t = 20)),
    axis.title.y = element_text(size = 24, margin = margin(r = 20)),
    axis.text = element_text(size = 20),
    strip.background = element_rect(fill = "#DEDEDE"))

ggsave("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/Data/Fitness/221028_dry_weight.pdf", dpi = 300, width = 12, height = 7)
ggsave("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/101722ave_larval_weight.pdf", dpi = 300, width = 12, height = 7)


weight.plot.box <- 
  ggplot(aes(x = corn_trait, y = weight_per_larvae, fill= treatment), data = fit.dat.wcr2) +
  scale_fill_manual(values = c("#C3BD10", "#137098"), labels = c("Cover crop no-till", "Traditional")) +
  geom_boxplot(width=0.75, position = position_dodge(width = .9)) + scale_x_discrete(labels = c("Bt corn", "non-Bt corn")) +
  facet_wrap(~colony, scales = "fixed",labeller = labeller(colony = supp.labs)) +
  theme_classic() + xlab("Corn type") +ylab("Average larval weight (g)") + labs(fill = "Treatment") +
  theme(
    strip.text.x = element_text(size = 26),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 20),
    axis.title.x = element_text(size = 24, margin = margin(t = 20)),
    axis.title.y = element_text(size = 24, margin = margin(r = 20)),
    axis.text = element_text(size = 20),
    strip.background = element_rect(fill = "#DEDEDE"))

ggsave("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/Data/Fitness/221107_dry_weight_box_flip.pdf", dpi = 300, width = 12, height = 7)

wp2.m <- weight.plot2 + 
  geom_point(aes(x = corn_trait, y = mean_weight), data = sum.weight, position = position_dodge(.9)) +
  geom_errorbar(aes(ymin=mean_weight-sd, ymax=mean_weight+sd), data = sum.weight, width=0.07, position = position_dodge(0.5))

wp3 <- ggplot(rhizo_div_full, aes(x=corn_trait, y = Observed, fill = treatment)) +
  geom_violin() + scale_fill_manual(values=c("#C3BD10", "#137098"), labels = c("Cover crop", "Traditional")) +
  geom_boxplot(width=0.1, position = position_dodge(width = .9)) +
 geom_point( position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.2), alpha = .5) + labs(fill = "Treatment") + xlab(label = "Corn trait")
```


Let's look at survival data. 

In previous work, we found no significant difference between using a Poisson vs a negative binomial distribution. However, the model seemed to be overly complex. I think we should investigate other models.

```{r}
prop.fit.dat <- fit.dat.wcr %>% 
  mutate(prop_surv = larv_surv/6)

# summary of average weight and standard error
sum.surv <- prop.fit.dat %>% 
  dplyr::group_by(colony, corn_trait, treatment) %>% 
  dplyr::summarise(mean_surv = mean(prop_surv, na.rm = TRUE),
                   sd = std.error(prop_surv, na.rm = TRUE))

# generate figure with means, std. error and raw data

prop.surv.plot <- sum.surv %>% 
  ggplot(aes(x = corn_trait, y = mean_surv, colour= treatment)) +
  geom_point(position = position_dodge(0.5)) + 
  geom_errorbar(aes(ymin=mean_surv-sd, ymax=mean_surv+sd), width=.2, position = position_dodge(0.5)) +  facet_wrap(~colony, scales = "free_y")
prop.surv.plot + geom_point(aes(x = corn_trait, y = prop_surv), data = prop.fit.dat, position = position_dodge(0.5), alpha=0.5)

```

It looks like there may be some treatment effect. Let's analyze them both ways and see.

```{r}
## Model for survival
nlmm.mod.fit <- glmer.nb(larv_surv ~ treatment * corn_trait * colony + (1|block:corn_trait) + (1|block:colony),
                data = fit.dat.wcr, verbose = TRUE)

summary(nlmm.mod.fit)
emm <- emmeans(nlmm.mod.fit, ~ treatment*colony*corn_trait)
emm2 <- as.data.frame(emm)
emm.cl <- emmeans(nlmm.mod.fit, ~ colony)
emm.cl2 <- as.data.frame(emm.cl)

emm.plot <- emm2 %>% 
  ggplot(aes(x = corn_trait, y = emmean, colour= treatment)) +
  geom_point(position = position_dodge(0.5)) + 
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), width=.2, position = position_dodge(0.5)) +  facet_wrap(~colony, scales = "free_y")

emmcl.plot <- emm.cl2 %>% 
  ggplot(aes(x = colony, y = emmean)) +
  geom_point(position = position_dodge(0.5)) + 
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), width=.2, position = position_dodge(0.5)) 


nlmm.mod.fit2 <- glmer(larv_surv ~ treatment * corn_trait * colony + (1|block:corn_trait) + (1|block:colony),
                      data = fit.dat.wcr, family = poisson)

summary(nlmm.mod.fit2)
```
We find no differences in treatment. However, the mortality data shows that Bt is still effective against susceptible insects in both cases.

We can also check hcw to estimate fitness. Previous analysis found that the residuals are normally distributed.

```{r}
aov.fit.mod.hcw <- aov(HCW_20 ~ treatment * corn_trait * colony + block + block*colony + block*corn_trait, data = fit.dat.wcr)
shapiro.test(residuals(aov.fit.mod.hcw))
summary(aov.fit.mod.hcw)

```


Let's look at log response ratios of the dry weight to see if there is a change in the effectiness of Bt in the treatments.

```{r}
relval.dat <- read_csv("/Users/PaddockK/Documents/Projects/Cover_crop/cover_crop_git/Data/Fitness/220302_relative_fitness.csv")



```


